---
title: "p8105_hw2_hq2182"
author: "Hanfei Qi"
date: "9/24/2020"
output: html_document
---

```{r setup, message = FALSE, warning = FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1
  
First, define a path to the dataset. 

```{r}
path_to_data_pb1 = "./Trash-Wheel-Collection-Totals-8-6-19.xlsx"
```
  
Read and clean the Mr. Trashwheel dataset. 

```{r}
trashwheel_df = 
	read_xlsx(
		path = path_to_data_pb1,
		sheet = "Mr. Trash Wheel",
		range = cell_cols("A:N")) %>% 
	janitor::clean_names() %>% 
	drop_na(dumpster) %>% 
	mutate(
		sports_balls = round(sports_balls),
		sports_balls = as.integer(sports_balls)
	)
```
  
Read precipitation data! For 2018 and 2017. 

```{r, message = FALSE, warning = FALSE}
precip_2018 = 
	read_excel(
		"./Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "2018 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2018) %>% 
	relocate(year)

precip_2017 = 
	read_excel(
		"./Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "2017 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2017) %>% 
	relocate(year)
```
  
Now combine annual precipitation.

```{r}
month_df = 
	tibble(
		month = 1:12,
		month_name = month.name
	)

precip_df = 
	bind_rows(precip_2018, precip_2017)

join_precip = left_join(precip_df, month_df, by = "month")
join_precip
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data from 2017 and 2018, with variables: `r colnames(precip_2018)`.There are a total of `r nrow(join_precip)` rows in our final precipation dataset, with variables: `r colnames(join_precip)`. Total precipitation in 2018 is `r sum(pull(precip_2018, total))`, while there is `r sum(pull(precip_2017, total))` in 2017.

  

## Problem 2

Read and clean NYC Transit data.

```{r, message = FALSE}
subway_df = 
  read_csv(
    file = "./NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_names = TRUE) %>%
  janitor::clean_names() %>%
  select(c(line:entry, vending, ada)) %>% 
  mutate(entry = recode(
                        entry, 
                        `YES` = TRUE,
                        `NO` = FALSE))
```

The NYC Transit data originally includes `r nrow(subway_df)` rows, with variables: company, line, station name, station latitude/longitude, route number, entrance type, entry/exit only, vending, staffing, staff hour, ada compliant, free crossover. It also includes the station's nearby streets and entrance location.

The final data contains variables: `r colnames(subway_df)`, with `r dim(subway_df)` dimension (row, column). Additionally, the "entry" variable was converted from character to logical. These data are not tidy because variables from a category are separated into different columns such as route1,2,3 ...

  
Now, find distinct stations, stations with ADA compliant, and station without vending allow entrance.

```{r}
distinct_station = distinct(
  subway_df, 
  station_name, line,
  .keep_all = TRUE
)

station_ada = subset(distinct_station, ada == TRUE)
station_no_vending = subset(distinct_station, vending == "NO")
```

There are `r nrow(distinct_station)` distinct stations, `r nrow(station_ada)` of them are ADA compliant, `r nrow(station_no_vending) / nrow(distinct_station) * 100`% of their entrances/exit without vending allow entrance.


  
Reformat data so that route number and route name are distinct variables.

```{r}
reformat_route = distinct_station %>%
  mutate(
    route8 = as.character(route8),
    route9 = as.character(route9),
    route10 = as.character(route10),
    route11 = as.character(route11)) %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_name",
    values_to = "route_number"
  )

reformat_distinct_station = distinct(
  reformat_route, 
  station_name, line,
  .keep_all = TRUE
)

with_number_A = subset(
  reformat_distinct_station, 
  route_number == "A")

ada_compliant = subset(
  with_number_A,
  ada == TRUE
)
```

There are `r nrow(reformat_distinct_station)` distinct station, which is the same number as I observed before. `r nrow(with_number_A)` distinct stations serve the A train, `r nrow(ada_compliant)` of them are ADA compliant.


## Problem 3
  
Define path to three datesets.

```{r}
path_to_pols_data = "./fivethirtyeight_datasets/pols-month.csv"
path_to_snp_data = "./fivethirtyeight_datasets/snp.csv"
path_to_unemp_data = "./fivethirtyeight_datasets/unemployment.csv"
```
  
Read and clean data in pols-month.csv;  
break variable mon into yaer, month, and day;  
replace month number with month name;  
creat variable president with values gop and dem;  
remove variables prez_dem and prez_gop;  
remove day variable.

```{r pols data, message = FALSE, warning = FALSE}
pols_df = read_csv(
  file = path_to_pols_data
) %>% 
  janitor::clean_names() %>% 
  separate(col = mon, 
           into = c("year", "month_number", "day")) %>% 
  mutate(
    year = as.integer(year),
    month_number = as.integer(month_number),
    day = as.integer(day)
  )

lower_month_name = month.name %>% 
  tolower()

month_df_q3 = 
	tibble(
		month_number = 1:12,
		month = tolower(lower_month_name)
	)

pols_reform = 
  left_join(pols_df, month_df_q3, by = "month_number") %>% 
  select(-month_number) %>% 
  mutate(
    president = case_when(
      prez_dem == 1 ~ "dem",
      prez_gop == 1 ~ "gop",
    )) %>% 
  select(-prez_dem, -prez_gop, -day) %>% 
  relocate(year, month)
```
  
Read, clean, and arrange data in snp.csv;
move variables year and month to be the leading columns.

```{r snp data, message = FALSE, warning = FALSE}
snp_df = read_csv(
  file = path_to_snp_data
) %>% 
  janitor::clean_names() %>% 
  separate(col = date, 
           into = c("month_number", "day", "year")) %>% 
  mutate(
    year = as.integer(year),
    month_number = as.integer(month_number),
    day = as.integer(day)
  )

snp_reform = 
  left_join(snp_df, month_df_q3, by = "month_number") %>% 
  select(-month_number, -day) %>% 
  relocate(year, month)
```
  
Read, clean, and tidy data in unemployment.csv;
variables Jan:Dec are values in a new variable month_name;
values of each month are values in a new variable unemploy_percent;
move variables year and month to be the leading columns.

```{r unemployment data, message = FALSE, warning = FALSE}
unemp_df = read_csv(
  file = path_to_unemp_data
) %>% 
  janitor::clean_names() %>%
  pivot_longer(
    jan:dec,
    names_to = "month_name",
    values_to = "unemploy_percent"
  )

month_df_unemp = 
	tibble(
		month_name = c("jan", "feb", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec"),
		month = tolower(lower_month_name))

unemp_reform = 
  left_join(unemp_df, month_df_unemp, by = "month_name") %>% 
  select(-month_name) %>% 
  relocate(year, month)
```
  
Join the datasets by merging snp into pols, and merging unemployment into the result.

```{r}
q3_final_df = 
  left_join(pols_reform, snp_reform, by = c("month", "year")) %>% 
  left_join(unemp_reform, by = c("month", "year"))
```

There are three original datasets.  
The pols-month.csv includes `r nrow(pols_df)` observations of `r ncol(pols_df)` variables. It's related to the number of national politicians who are democratic or republican at any given time.  
The snp.csv includes `r nrow(snp_df)` observations of `r ncol(snp_df)` variables. It's related to S&P stock index at coordinating time points. S&P represents measure of stock market as a whole.    
The unemployment.csv includes `r nrow(unemp_df)` observations of `r ncol(unemp_df)` variables. It represents percentage of unemployment in certain time.  

The data arranging steps are described before each code chunk.  
In conclusion, the final data is the result of left_join three reformed datasets by using keys "month" and "year".  There are `r nrow(q3_final_df)` observations of `r ncol(q3_final_df)` variables in the final dataset. The range of year is `r range(q3_final_df$year)`. The final data contains variables: `r colnames(q3_final_df)`.  


Here is the list of variables in the final data:

* year: year of the observation
* month: month of the observation
* gov_gop: the number of republican governors on the associated date
* sen_gop: the number of republican senators on the associated date
* rep_gop: the number of republican representatives on the associated date
* gov_dem: the number of democratic governors on the associated date
* sen_dem: the number of democratic senators on the associated date
* rep_dem: the number of democratic representatives on the associated date
* president: indicator of whether the president was republican on the associated date (gop = republican, dem = democratic)
  * Some values are N/A because of historical events
* close: the closing values of the S&P stock index on the associated date
* unemply_percent: percentage of unemployment in associated time