---
title: "p8105_hw2_hq2182"
author: "Hanfei Qi"
date: "9/24/2020"
output: html_document
---

```{r setup, message = FALSE, warning = FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1

First, define a path to the dataset. 

```{r}
path_to_data_pb1 = "./Trash-Wheel-Collection-Totals-8-6-19.xlsx"
```

Read the Mr. Trashwheel dataset. 

```{r}
trashwheel_df = 
	read_xlsx(
		path = path_to_data_pb1,
		sheet = "Mr. Trash Wheel",
		range = cell_cols("A:N")) %>% 
	janitor::clean_names() %>% 
	drop_na(dumpster) %>% 
	mutate(
		sports_balls = round(sports_balls),
		sports_balls = as.integer(sports_balls)
	)
```

Read precipitation data! For 2018 and 2017. 

```{r, message = FALSE, warning = FALSE}
precip_2018 = 
	read_excel(
		"./Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "2018 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2018) %>% 
	relocate(year)

precip_2017 = 
	read_excel(
		"./Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "2017 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2017) %>% 
	relocate(year)
```

Now combine annual precipitation.

```{r}
month_df = 
	tibble(
		month = 1:12,
		month_name = month.name
	)

precip_df = 
	bind_rows(precip_2018, precip_2017)

left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Addtional data sheets include month precipitation data.

## Problem 2


```{r, message = FALSE}
subway_df = 
  read_csv(
    file = "./NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_names = TRUE) %>%
  janitor::clean_names() %>%
  select(c(line:entry, vending, ada)) %>% 
  mutate(entry = recode(
                        entry, 
                        `YES` = TRUE,
                        `NO` = FALSE))
```

TO DO: Write a short paragraph about this dataset – explain briefly what variables the dataset contains, describe your data cleaning steps so far, and give the dimension (rows x columns) of the resulting dataset. Are these data tidy?

```{r}
distinct_station = distinct(
  subway_df, 
  station_name, line,
  .keep_all = TRUE
)

station_ada = subset(distinct_station, ada == TRUE)
station_vending = subset(distinct_station, vending == "YES")
```

```{r}
reformat_route = distinct_station %>%
  mutate(
    route8 = as.character(route8),
    route9 = as.character(route9),
    route10 = as.character(route10),
    route11 = as.character(route11)) %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_name",
    values_to = "route_number"
  )

reformat_distinct_station = distinct(
  reformat_route, 
  station_name, line,
  .keep_all = TRUE
)

with_number_A = subset(
  reformat_distinct_station, 
  route_number == "A")

ada_compliant = subset(
  reformat_distinct_station,
  ada == TRUE
)
```

## Problem 3

```{r}
path_to_pols_data = "./fivethirtyeight_datasets/pols-month.csv"
path_to_snp_data = "./fivethirtyeight_datasets/snp.csv"
path_to_unemp_data = "./fivethirtyeight_datasets/unemployment.csv"
```

```{r pols data, message = FALSE, warning = FALSE}
pols_df = read_csv(
  file = path_to_pols_data
) %>% 
  janitor::clean_names() %>% 
  separate(col = mon, 
           into = c("year", "month_number", "day")) %>% 
  mutate(
    year = as.integer(year),
    month_number = as.integer(month_number),
    day = as.integer(day)
  )

lower_month_name = month.name %>% 
  tolower()

month_df_q3 = 
	tibble(
		month_number = 1:12,
		month = tolower(lower_month_name)
	)

pols_reform = 
  left_join(pols_df, month_df_q3, by = "month_number") %>% 
  select(-month_number) %>% 
  mutate(
    president = prez_gop + gov_gop + sen_gop + rep_gop + prez_dem + gov_dem  + sen_dem + rep_dem) %>% 
  select(-prez_dem, -prez_gop, -day) %>% 
  relocate(year, month)
```

```{r snp data, message = FALSE, warning = FALSE}
snp_df = read_csv(
  file = path_to_snp_data
) %>% 
  janitor::clean_names() %>% 
  separate(col = date, 
           into = c("month_number", "day", "year")) %>% 
  mutate(
    year = as.integer(year),
    month_number = as.integer(month_number),
    day = as.integer(day)
  )

snp_reform = 
  left_join(snp_df, month_df_q3, by = "month_number") %>% 
  select(-month_number, -day) %>% 
  relocate(year, month)
```

```{r unemployment data, message = FALSE, warning = FALSE}
unemp_df = read_csv(
  file = path_to_unemp_data
) %>% 
  janitor::clean_names() %>%
  pivot_longer(
    jan:dec,
    names_to = "month_name",
    values_to = "unemploy_percent"
  )

month_df_unemp = 
	tibble(
		month_name = c("jan", "feb", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec"),
		month = tolower(lower_month_name))

unemp_reform = 
  left_join(unemp_df, month_df_unemp, by = "month_name") %>% 
  select(-month_name) %>% 
  relocate(year, month)
```

```{r}
q3_final_df = 
  left_join(pols_reform, snp_reform, by = "month") %>% 
  left_join(unemp_reform, by = "month")
```

